Do return series have power-law tails with the same index?

### 2017-04-11 papers/FX/draft.tex
Merged Thomas' revision.

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
%\usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

\newcommand{\BM}{Brownian motion}
\newcommand{\BB}{Brownian bridge}
\newcommand{\con}{convergence}
\newcommand{\ev}{extreme value}
\newcommand{\evt}{extreme value theory}
\newcommand{\evd}{extreme value distribution}
\newcommand{\clp}{central limit problem}
\newcommand{\nc}{norming constant}
\newcommand{\st}{such that}
\newcommand{\fif}{if and only if}
\newcommand{\wrt}{with respect to}
\newcommand{\chf}{characteristic function}
\newcommand{\fct}{function}
\newcommand{\mdoa}{maximum domain of attraction}
\newcommand{\ds}{distribution}
\newcommand{\gev}{generalized extreme value distribution}
\newcommand{\gpd}{generalized Pareto distribution}
\newcommand{\pwm}{probability-weighted moments}
\newcommand{\mle}{maximum likelihood estimator}
\newcommand{\mel}{mean excess function}
\newcommand{\an}{asymptotically normal}
\newcommand{\ac}{absolutely continuous}
\newcommand{\rep}{representation}
\newcommand{\cmt}{continuous mapping theorem}
\newcommand{\seq}{sequence}
\newcommand{\pr} {{\bf Proof.}}
\newcommand{\ins}{insurance}
\newcommand{\mat}{mathematic}
\newcommand{\pro}{probabilit}
\newcommand{\pros}{probabilities}
\newcommand{\ms}{measure}
\newcommand{\mgf}{moment generating function}
\newcommand{\ld}{large deviation}
\newcommand{\bfx}{{\bf x}}
\newcommand{\bfX}{{\bf X}}
\newcommand{\bfB}{{\bf B}}
\newcommand{\bfY}{{\bf Y}}
\newcommand{\bfy}{{\bf y}}
\newcommand{\bfA}{{\bf A}}
\newcommand{\bfG}{{\bf G}}
\newcommand{\bfO}{{\bf 0}}
\newcommand{\bfZ}{{\bf Z}}
\newcommand{\bfz}{{\bf z}}
\newcommand{\bfS}{{\bf S}}
\newcommand{\bfa}{{\bf a}}
\newcommand{\bfb}{{\bf b}}
\newcommand{\bfe}{{\bf e}}
\newcommand{\bft}{{\bf t}}
\newcommand{\bfI}{{\bf I}}
\newcommand{\bfs}{{\bf s}}
\newcommand{\bfh}{{\bf h}}
\newcommand{\bfunit}{{\bf h}}
\newcommand{\ARCH}{{\rm ARCH}}
\newcommand{\E }{{\mathbb E}}
\renewcommand{\P }{{\mathbb P}}
\newcommand{\Q }{{\mathbb Q}}
\newcommand{\1}{{\mathbf 1}}

@@

\allowdisplaybreaks
\begin{document}
\title{Tail-indices and scale parameters of returns series}
\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei, }
\date{\today}

@@

in the literature how heavy tails really are. For example, Barndorff-Nielsen and Shephard
\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor
``semi-heavy'' tails, i.e., tails which are comparable with those of a
gamma \ds . On the other hand, tails of returns have been studied in
great detail in the extreme value community. Among extreme value
specialists there is general agreement that returns $X_t$ have tails
of power-law-type, i.e.,

@@

\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad
\P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,

@@

where $c_{\pm}$ are positive constants.\footnote{Here and in what

@@

al. \cite{Embrechts1997}, Jansen and de Vries

@@

In the extreme value literature it is common to replace the constants $c_\pm$ by a suitable

@@

al. \cite{Embrechts1997}, Chapter 3. In this

@@

of data are believed to be infinite. Tails of the type \eqref{eq:1}

@@

Fr\'echet \ds\ $\Phi_\alpha(x)=\exp(-x^{-\alpha})$ for $x>0$. Equivalently, power-law tails are prescribed by the 
generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e.,

@@

\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha)^{-\alpha}\,,\qquad u\to\infty\,.

@@

are typically found in the range $(2.5,4)$. For an illustration, see

@@

in three sectors of the Standard \& Poor 500 index  are shown. The

@@

data from 1 January 2010 to 1 January 2015.

@@

The estimator of the tail-index $\alpha$  which favored in the literature  is the {\em Hill estimator}; the graphs in 

@@

that the simulateneous estimation of the tail-index and the scale parameter are strongly related.

@@

return data. In particular, we address the power-law tails of univariate and multivariate GARCH models as potential models

@@

have power-tails due to their relation with so-called {\em \sre s}.

@@

is an extension of the concept of Disappointment Aversion (DA) of Gul
 \cite{gul1991theory}. The Gul paper derives the DA from first principles (axiomatic).

@@

\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
In this section, we assume the model \eqref{eq:1} for the tails of the marginal
\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
that this series constitutes a strictly stationary \seq . In what follows, we focus
on the left tail of the \ds , i.e., on the losses. However, it is common to present the tail index estimators
for positive data. Therefore we will multiply the losses by minus one, swapping the negative with the positive values.
 The two parameters -- the tail-index  $\alpha_{\rm low}$ and the scale parameter
$c_-$ -- are assumed positive. They play crucial roles for the understanding the risk hidden in the data, hence 
for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
useful description of the risk. Alternatively, these parameters can be
used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
%With these motives in mind, we present a survey of the
%tail indices and scale parameters of 3 sectors, namely ``Energy'',
%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
%index.

@@

Various estimators of the tail index $\alpha_{\rm low}$  in the model
\eqref{eq:1} have been proposed in the literature;
see Embrechts et al. \cite{Embrechts1997}, de
Haan and Ferreira \cite{dehaan:ferreira:2006}, Resnick

@@

Given a sample $X_1,\ldots,X_n$ from \eqref{eq:1}, one can calculate

@@

  \hat \alpha_{\rm low} = \Big(

@@

for the estimation. The estimator $\hat \alpha_{\rm low}$ is 

@@

\P(X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\quad x>K\,,

@@

dependent \seq s and de Haan and Ferreira provide

@@

of the \ds\ of the $X_t$. This leads to bias
of the estimator. On the other hand, by construction $\hat \alpha_{\rm
  low}$ is an average of $k$ log-differences of the 

@@

fact does not make the estimation of $\alpha_{\rm low}$ an  

@@

a so-called {\em Hill plot} is recommended where $\hat \alpha$ is

@@

region in the plot where $\hat \alpha_{\rm low}$ is relatively

@@

see Embrechts et al. \cite{Embrechts1997},

@@

\cite{drees:resnick:2000}. In Figure~\ref{fig:1},
we exhibit 95\% \asy\ confidence bands which is based on the \clt

@@

\sqrt k\, \big(\hat \alpha_{\rm low} - \alpha_{\rm low}\big) \std N(0, \alpha^2)\,,

@@

i.e., $\hat \alpha_{\rm low}$ is \asy ally unbiased and has variance

@@

$\sqrt{n}$-rates. This fact is one explanation for the fact that it is

@@

$\alpha_{\rm low}$. There exist various other reasons why one should

@@

al. \cite{Embrechts1997} for illustrations 

@@

of the data and to dependence. For example, results in Drees and

@@

confidence bands could be evaluated.

@@

  \caption{\small Hill estimates of lower tail indices of
    daily return series in sectors of Standard \& Poor
    500 index. The figures from top to bottom are ``Energy'',

@@

    Each circle coresponds to a Hill estimate; the grey

@@

    of its approximate normal distribution.
    The low and high red lines mark the median of the 2.5\% quantiles
    and median of the 97.5\% quantiles, respectively.
    The data are from {\it Yahoo Finance} and the labels on
    the horizontal axis are Yahoo symbols of the stocks. The data span
    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
    observations. $k=50$ lower order statistics are used
    for computing $\alpha_n$. 

@@

In Figure~\ref{fig:1} we see that confidence intervals of the Hill
estimates of the losses in the ``Energy'' and ``Consumer Staples''
sectors of the S\& P 500 index, as well as those of a 
large portion of losses in the ``Information Technology'' sector,
overlap. This fact indicates that the returns in each sector may 
have comparable tail indices.

Hoga's test on the change of extreme quantiles
may provide some insight about how similar these tail indices are,
because different tail indices are likely to result in different
extreme quantiles. Nevertheless, changes in the extreme quantiles can
be the result of changed tail indices but can also be the result of
changed scale parameters. Thus we first investigate the scale
parameters of stocks in the same sectors before applying Hoga's test.

\subsection{Hill estimates of lower-tail scale parameters}
\label{sec:HillScaleEstimates}

For convenience, we denote the scale parameter by $A$.
%% It is the constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
Hill \cite{hill1975simple} put forward an estimator of $A$ using his
tail index  estimator:
\[
\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
\]
where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
sequence $X_t$ as defined in section\ref{sec:Hill}. $\hat \alpha$
is the Hill tail index estimator.
Estimates of $A$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in figure
\ref{fig:sectors_parameters}, along with estimates obtained
using the hybrid method described in section
\ref{sec:hybrid_estimation}.

%% Hence it is informative to test
%% the hypothesis that the tail indices of these return series are equal.
%% Hoga \cite{Hoga2017} devised tests suitable this purpose. We adopt his
%% approach but use our own implementation. The next section has more
%% details.

@@

\subsection{Hoga's Test}
\label{sec:Hoga}

Hoga \cite{Hoga2017} proposed his test for a time series
$\{X_i\}_{i=1}^n$ with tail functions
$\bar F_i(\cdot)$, i.e. $\P(X_i > x) = \bar F_i(x)$. The null
hypothesis is
\begin{hypothesis}
  $\mathcal H_0$: For a given $p$ and all $1 \leq i < j \leq n$,
  $F_i^{-1}(p) = F_j^{-1}(p)$
\end{hypothesis}
while the alternative is $\exists t^* \in [t_0, 1 - t_0]$ such that
\[
\bar F_1^{-1}(p) = \bar F_2^{-1}(p) = \cdots =
\bar F_{\floor{t^* n}}^{-1}(p) \neq
\bar F_{\ceil{t^* n}}^{-1}(p) = \cdots = F_n^{-1}(p)
\]
where $t_0 \in (0, 1/2)$ and $\hat x_p(s, t)$ is the Weissman
estimator applied to the partial series
$X_{\ceil{ns}}, \dots, X_{\floor{nt}}$ using $\ceil{k(t - s)}$
upper-order statistics. See Embrechts, Kl\"uppelberg and T.Mikosch
\cite{Embrechts1997}, p.347.

Hoga proposed the following test statistics:
\[
Q_1 = \sup_{t \in [t_0, 1 - t_0]}
\left\{
  {
    \left[
      t (1 - t) \ln\left(
      \hat x_p(0, t)
      \over
      \hat x_p(t, 1)
      \right)
    \right]^2
    \over
    \int_{t_0}^t
    \left[
      s \ln \left(
      \hat x_p(0, s)
      \over
      \hat x_p(0, t)
      \right)
    \right]^2 ds
    +
    \int_{t}^{1 - t_0}
    \left[
      (1 - s) \ln \left(
      \hat x_p(s, 1)
      \over
      \hat x_p(t, 1)
      \right)
    \right]^2 ds
  }
\right\}
\]
Under the null hypothesis and suitable dependence conditions, Hoga
showed that as $n \to \infty$,
\begin{equation}
  \label{eq:Q1_distr}
  Q_1 \overset{d}{\to}
  \sup_{t \in [t_0, 1 - t_0]} q_t
\end{equation}
where
\[
q_t = 
  {
    [W(t) - t W(1)]^2
    \over
    \int_{t_0}^t [W(s) - {s \over t}W(t)]^2 ds
    +
    \int_{t}^{1 - t_0}
    \left[
      W(1) - W(s) - {1 - s \over 1 - t}
      (W(1) - W(t))
      \right]^2 ds
  }
\]
$W(t)$ with $t \in (0, 1)$ is a standard Brownian motion.

When applied to our problem, we would like to test, first of all,
whether there is a change of tail index in each of two 
series and, if there is no change in either series, whether the two
series have the same tail index. Obviously, for either question, if
the tail index changes, the extreme quantile will change.
For the first question, Hoga's test is directly applicable; for the
second, we concatenate the two series before applying the
test.

Figure \ref{fig:Hoga_Single} shows the test statistics of
stocks in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
500 index. The null hypothesis here is that the extreme quantile
remains the same througout the selected period of time.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Energy_Single.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Consumer_Staples_Single.pdf}
  \end{minipage}
  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
    lines are, respectively the 85\%, 90\% and 95\% quantiles of the
    distribution of \eqref{eq:Q1_distr}}.
  \label{fig:Hoga_Single}
\end{figure}

To test whether two series share the same tail index, we concatenate
the two series and apply the Hoga test. For the ``Energy'' and the
``Consumer Staples'' sector, we summarise the results in figure
\ref{fig:Hoga_pair}.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 1cm, 1cm, 1cm}, clip
    ]{Hoga_Energy_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 1cm, 1cm, 1cm}, clip
    ]{Hoga_Consumer_Staples_pair.pdf}
  \end{minipage}
  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of
    S\&P 500. The green, blue and red colors represent, respectively,
    exceedances of the 85\%, 90\% and 95\% quantiles of the distribution of
    \eqref{eq:Q1_distr}.
    Grey points stand for pairs for which the test statistic is below
    the 85\% quantile;
    black points represent pairs for which
    computation of the test statistics fails for given precision
    requirements and time limits.}
  \label{fig:Hoga_pair}
\end{figure}
Clearly, these figures show that the null
hypothesis of an equal tail index is rejected for more pairs in the
``Energy'' sector than it is for those in the ``Consumer Staples''
sector. This suggests that lower tail indices of stocks in the
``energy'' sector are more spread out than are those of the ``Consumer
Staples'' sector.

Also observe that while 3 stocks, say A, B and C, test in favor of
relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
happens that another test on B and C is supportive of $\alpha_B =
\alpha_C$ . This is due to the limited power of the test. Based
on such results, one may guess $\alpha_B$ lies between $\alpha_A$ and
$\alpha_C$. The test is unable to recognize the smaller differences
between $\alpha_A, \alpha_B$ and between $\alpha_B, \alpha_C$.

To have an idea about the power of Hoga's test, we run it on
concatenated samples drawn from t-distributions with different degrees
of freedom. Figure \ref{fig:t_sim_pair} shows the result.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth, trim={1.0cm 1.5cm 1cm 2cm},
  clip]{t_sim_pair.pdf}
  \caption{Test Statistics of concatenated t-samples with different
    degrees of freedom. Numbers on the axis are the degrees of freedom.}
  \label{fig:t_sim_pair}
\end{figure}
From the figure one can see that the power of the test decreases as
the minimum tail index of a concatenated pair increases.

>>>>

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
%\usepackage[bookmarks=true]{hyperref}
%\usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

\newcommand{\BM}{Brownian motion}
\newcommand{\BB}{Brownian bridge}
\newcommand{\con}{convergence}
\newcommand{\ev}{extreme value}
\newcommand{\evt}{extreme value theory}
\newcommand{\evd}{extreme value distribution}
\newcommand{\clp}{central limit problem}
\newcommand{\nc}{norming constant}
\newcommand{\st}{such that}
\newcommand{\fif}{if and only if}
\newcommand{\wrt}{with respect to}
\newcommand{\chf}{characteristic function}
\newcommand{\fct}{function}
\newcommand{\mdoa}{maximum domain of attraction}
\newcommand{\ds}{distribution}
\newcommand{\gev}{generalized extreme value distribution}
\newcommand{\gpd}{generalized Pareto distribution}
\newcommand{\pwm}{probability-weighted moments}
\newcommand{\mle}{maximum likelihood estimator}
\newcommand{\mel}{mean excess function}
\newcommand{\an}{asymptotically normal}
\newcommand{\ac}{absolutely continuous}
\newcommand{\rep}{representation}
\newcommand{\cmt}{continuous mapping theorem}
\newcommand{\seq}{sequence}
\newcommand{\pr} {{\bf Proof.}}
\newcommand{\ins}{insurance}
\newcommand{\mat}{mathematic}
\newcommand{\pro}{probabilit}
\newcommand{\pros}{probabilities}
\newcommand{\ms}{measure}
\newcommand{\mgf}{moment generating function}
\newcommand{\ld}{large deviation}
\newcommand{\bfx}{{\bf x}}
\newcommand{\bfX}{{\bf X}}
\newcommand{\bfW}{{\bf W}}
\newcommand{\bfB}{{\bf B}}
\newcommand{\bfY}{{\bf Y}}
\newcommand{\bfy}{{\bf y}}
\newcommand{\bfA}{{\bf A}}
\newcommand{\bfG}{{\bf G}}
\newcommand{\bfO}{{\bf 0}}
\newcommand{\bfZ}{{\bf Z}}
\newcommand{\bfz}{{\bf z}}
\newcommand{\bfS}{{\bf S}}
\newcommand{\bfa}{{\bf a}}
\newcommand{\bfb}{{\bf b}}
\newcommand{\bfe}{{\bf e}}
\newcommand{\bft}{{\bf t}}
\newcommand{\bfI}{{\bf I}}
\newcommand{\bfs}{{\bf s}}
\newcommand{\bfh}{{\bf h}}
\newcommand{\bfunit}{{\bf h}}
\newcommand{\ARCH}{{\rm ARCH}}
\newcommand{\E }{{\mathbb E}}
\renewcommand{\P }{{\mathbb P}}
\newcommand{\Q }{{\mathbb Q}}
\newcommand{\1}{{\mathbf 1}}

@@

\allowdisplaybreaks
\begin{document}
\title{Do return series have power-law tails with the same index?}
\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
\date{\today}

@@

in the literature about how heavy these tails really are. For example, Barndorff-Nielsen and Shephard
\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
which are comparable with those of a gamma \ds . On the other
hand, tails of returns have been studied in great detail 
in the extreme value community. Among extreme value specialists
there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,

@@

\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,

@@

where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive constants.\footnote{Here and in what

@@

al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries

@@

In the extreme value literature it is common to replace the constants $c_\pm$ by suitable

@@

al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this

@@

of data are believed to be infinite. Tails of type \eqref{eq:1}

@@

Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$ for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
upper tail described in \eqref{eq:1} converge in \ds\ to $\Phi_{\alpha_{\rm up}}$.
Equivalently, power-law tails are prescribed by the 
generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e., for a suitable positive scaling \fct\ $a(u)$,

@@

\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm up}}\,,\qquad u\to\infty\,.

@@

are typically found in the range $(2,4)$. For an illustration, see

@@

in three sectors of the Standard \& Poors 500 index  are shown. The

@@

data from 1 January 2010 to {\red why 1 January?} 1 January 2015.

@@

The estimator of the tail-index $\alpha>0$  in the model
\P(X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
favored in the literature  is the {\em Hill estimator}; the graphs in 

@@

that the simultaneous estimation of the tail-index and the scale parameter are strongly related.

@@

return data. In particular, in Section~\ref{subsec:garch} we address the power-law tails of univariate and multivariate GARCH models as potential models

@@

have power-law tails due to their relation with so-called {\em \sre s}. Moreover, some of the {\em standard} multivariate 
GARCH models as the CCC or BEKK ensure that the marginal \ds s have power-law tails with the same index.

@@

is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991}. The Gul paper derives the DA from first principles (axiomatic).

@@

\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
In this section, we assume the model \eqref{eq:1} for the tails of the marginal
\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
that this series constitutes a strictly stationary \seq . In what follows, we focus
on the left tail of the \ds , i.e., on the losses. 
However, it is common to present the tail index estimators
for positive data. Therefore we will multiply the losses $X_t$ by minus one, swapping the negative with the positive values.
For simplicity, we also suppress subscripts in the notation:
\beam\label{eq:1a}
\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
useful description of the risk. Alternatively, these parameters can be
used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
%With these motives in mind, we present a survey of the
%tail indices and scale parameters of 3 sectors, namely ``Energy'',
%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
%index.

@@

Various estimators of the tail index $\alpha$  in the model
\eqref{eq:1a} have been proposed in the literature;
see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
Haan and Ferreira \cite{haan:ferreira:2006}, Resnick

@@

Given a sample $-X_1,\ldots,-X_n$ whose marginal \ds\ satisfies  \eqref{eq:1a}, calculate

@@

  \hat \alpha_k = \Big(

@@

for the estimation. The estimator $\hat \alpha_k$ is 

@@

\beam\label{eq:3}
\P(-X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\qquad x>K\,,

@@

dependent \seq s and de Haan and Ferreira \cite{haan:ferreira:2006} provide

@@

of the \ds\ of the $-X_t$, leading to a bias
of the estimator. On the other hand, by construction, $\hat \alpha_k$ is an average of $k$ log-differences of the 

@@

fact does not make the estimation of $\alpha$ an  

@@

a so-called {\em Hill plot} is recommended where $\hat \alpha_k$ is

@@

region in the plot where $\hat \alpha_k$ is relatively

@@

see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},

@@

\cite{drees:resnick:2000}. 
In Figure~\ref{fig:1},
we exhibit 95\% \asy\ confidence bands derived from  the \clt

@@

\sqrt k\, \big(\hat \alpha_k - \alpha\big) \std N(0, \alpha^2)\,,

@@

i.e., $\hat \alpha$ is \asy ally unbiased and has variance

@@

$1/\sqrt{n}$-rates. This fact is one explanation for the fact that it is

@@

$\alpha$. There exist various other reasons why one should

@@

al. \cite{embrechts:klueppelberg:mikosch:1997} and Resnick \cite{resnick:1987} for illustrations 

@@

of the data and to dependence. For example, results in Drees \cite{drees:2008}, and Drees and

@@

confidence bands can be evaluated and therefore the bands shown in Figure~\ref{fig:1}
just show some benchmark which holds in the iid case and under additional conditions on the tail \asy s.

@@

  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail indices $\alpha$ of
    daily return series in sectors of the S\&P 500
    index. The data span
    from 2010-01-01 to 2015-01-01 and comprise $n=1304$ observations.
The graphs from top to bottom correspond to the ``Energy'',

@@

    Each circle corresponds to a Hill estimate $\hat\alpha_k$; the grey

@@

    of its approximate normal distribution; see \eqref{eq:2} and the discussion following it for an 
interpretation.
    The lower and upper red lines mark the medians of the 2.5\% 
    and 97.5\% quantiles, respectively, evaluated from all stocks in the sector.
    The data are taken from {\it Yahoo Finance}; the labels on
    the horizontal axes are Yahoo symbols of the stocks. 

@@

In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
estimates of the losses in the ``Energy'' and ``Consumer Staples''
sectors of the S\&P 500 index, as well as those of a 
large portion of losses in the ``Information Technology'' sector.
This fact indicates that the returns in each sector may 
have comparable tail indices.
\par
Hoga's \cite{hoga:2016} test about the change of extreme quantiles
may provide some further insight about how similar these tail indices are.
Different tail indices are likely to result in different
extreme quantiles. Nevertheless, changes in the extreme quantiles may also
result from changing scale parameters in the tail. Therefore  we first investigate the scale
parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$. 
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = {k \over n} X_{(n-k+1)}\,.
\eeao
Estimates of $K$ {\red or $K^\alpha$?} of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. {\blue The results are shown in Figure~\ref{fig:sectors_parameters} {\red ???} along with estimates obtained
using the hybrid method described in section
\ref{sec:hybrid_estimation}.} {\red Figures of the Hill estimates of $K$ or $K^\alpha$ with Hill confidence bands. 
Discuss confidence bands and what you see in the data.}

%% Hence it is informative to test
%% the hypothesis that the tail indices of these return series are equal.
%% Hoga \cite{hoga:2016} devised tests suitable this purpose. We adopt his
%% approach but use our own implementation. The next section has more
%% details.

@@

\subsection{Tests for a change in the extreme tail}\label{sec:Hoga}
To the best of our knowledge, there do not exist statistical tests for comparing the 
tail indices or scale parameters of two or more dependent \ts\ in  a direct way. {\red construct a test for this}
We borrow a test from a recent paper by Hoga \cite{hoga:2016}. 
This test has been developed for a different kind of problem. Given a strictly stationary \ts\ $X_1,\ldots,X_n$
with a marginal \ds\ $F$ with right power-law tail, the goal is to test whether there is a structural break 
of the {\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to zero. If the tail index
{\em or} the scale parameter in a \ds\ of type \eqref{eq:3} change inside a sample, then it is likely
that the extreme quantiles change as well. We will test for a change of tail index or scale parameter 
in this indirect way.
The null hypothesis of the test in \cite{hoga:2016} is that there is no change of the extreme quantiles $F^{-1}(1-p)$ 
for $p=p_n\to 0$ in any subsample with indices
$t\in (n\,t_0,n(1-t_0))$ where $t_0$ is a fixed number in $(0,0.5)$. Writing $\hat x_p(a,b)$ for an estimator of the extreme $(1-p)$-quantile 
based on the subsample with indices $t\in (na,nb)$, the test statistic is given by
\beam\label{eq:4}
T_n = \sup_{s \in [t_0, 1 - t_0]}
  \dfrac{  \big[s (1 - s) \log \big(\hat x_p(0, s)/\hat x_p(s, 1)\big)
    \big]^2}{
    \int_{t_0}^s\big[r \log \big( \hat x_p(0, r)/\hat x_p(0, s)
      \big)
    \big]^2 dr
    +
    \int_{s}^{1 - t_0}
    \big[
      (1 - r) \log \big(
      \hat x_p(r, 1)/
      \hat x_p(s, 1)
      \big)
    \big]^2 dr}\nonumber\\
Under the null hypothesis, $(T_n)$ converges to a complicated \fct al of \BM\ on $[0,1]$; the \asy\ quantiles need to be 
evaluated by simulation.  
When applied to our problem we would like to test 
whether there is a change of the tail index {\em or} scale parameter in \eqref{eq:3} in each of the S\&P 500 
series in the distinct sectors. We also  want to get some indication about a possible change of tail index or
scale parameter from one series to another within a given sector. For this reason, we choose any pair of series
within a sector and concatenate each of the paired series. Then we run the test on the concatenated series.
Of course, despite the fact that we test changes of tail index or scale parameter 
{\em in a very indirect way} -- there may be many other reasons for the change of extreme quantiles in a sample -- 
we also concatenate two rather distinct series. Even if we assume that the two series come from related models
(such as GARCH), the parameters of these models will in general not be the same. Moreover, the concatenation
of two strictly stationary \ts\ is in general not strictly stationary. Therefore we have to be careful
with interpretations of the results of the tests.
In Figure~\ref{fig:Hoga_Single} we show the values of the test statistic $T_n$ (horizontal bars)
for $t_0=0.1$ and daily return series of stock in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
500 index. The ``null hypothesis'' is that the tail index and scale parameter remain the
same througout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
This fact may be an indication that the \ds\ inside a series is rather homogeneous.
Alternatively, it may show that the power of the test is very low. A possible reason for this suspicion is that the \con\ rate
of $(T_n)$ to its limit is very slow, i.e., the \asy\ \ds\ is not representative for the \ds\ of $T_n$ for the chosen $n$; for some
simulation evidence, see below.
To check whether any two series share the same tail index and scale parameter we concatenate
any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:Hoga_pair}.
These graphs show that the ``null
hypothesis'' of an equal tail index and scale parameter is rejected for more pairs in the
``Energy'' sector than it is for those in the ``Consumer Staples''
sector. This suggests that lower tail indices of stocks in the
``Energy'' sector are more spread out than those of the ``Consumer
Staples'' sector. Also observe that while 3 stocks, say A, B and C, test in favor of
the relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
happens that another test on B and C is supportive of $\alpha_B = \alpha_C$. Again,
this is due to the limited power of the test. Based
on such results, one may guess that $\alpha_B$ lies between $\alpha_A$ and
$\alpha_C$. The test is unable to recognize the smaller differences
between $\alpha_A, \alpha_B$ on one hand and between $\alpha_B, \alpha_C$ on the other hand.

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Energy_Single.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Consumer_Staples_Single.pdf}
  \end{minipage}
  \caption{Test statistic $T_n$ from \eqref{eq:4} for the  stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
    lines correspond to the 85\%, 90\% and 95\% quantiles of the limit \ds\ of $T_n$.  They are derived
by simulations from the limit \ds .}
  \label{fig:Hoga_Single}
\end{figure}

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 1cm, 1cm, 1cm}, clip
    ]{Hoga_Energy_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 1cm, 1cm, 1cm}, clip
    ]{Hoga_Consumer_Staples_pair.pdf}
  \end{minipage}
  \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
    ``Consumer Staples'' sectors of
    S\&P 500. The green, blue and red colors represent, respectively, the cases when  
    the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
    Grey points stand for pairs for which $T_n$ is below
    the 85\% \asy\ quantile;
    black points represent pairs for which
    the computation of $T_n$ fails for given precision
    requirements and time limits.}
  \label{fig:Hoga_pair}
\end{figure}
